---
title: "Homework 2"
author: Maya Bunyan
output: github_document
---

This is my solution to Homework 2! 

```{r load_packages}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read and clean 2017 and 2018 precipitation data.

```{r}
precip_2018 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1,
    ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1,
    ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017) 

precip_final = 
  left_join(precip_df, month_df, by = "month")
```


```{r, results='hide'}
#Calculate median sports balls for 2017 to include in inline code
trashwheel_2017 = 
  subset(trashwheel_df, year == 2017) 
  
sportsballs_2017 =
  median(pull(trashwheel_2017, sports_balls))
```


This dataset includes data collected by Mr. Trash Wheel, a water-wheel trash collector, in Baltimore, Maryland. Mr. Trash Wheel collects trash that flows into the inner harbor and then stores that trash in a dumpster, which is later taken to waste-to-energy plants and incinerated to create electricity. The dataset itself includes information on the date, dumpster number, as well as the amount and type of trash. There are `r nrow(trashwheel_df)` observations in our final dataset. We also saw a median number of  `r sportsballs_2017` sports balls in a dumpster in 2017. Other sheets include information on the precipitation levels in inches by month. In the combined 2017 and 2018 precipitation dataframe, there are `r nrow(precip_final)` observations, with a total precipitation level of `r sum(pull(precip_2018, total))` inches in 2018 and `r sum(pull(precip_2017, total))` inches in 2017.
